\documentclass[11pt]{tibop-article}
\TIBauthor[Gerner et al.]{Jeremias Gerner\affOne\and
            Klaus Bogenberger\affTwo\lastand \\
            Stefanie Schmidtner\affOne
          }
\TIBaffiliations{\affOne Technische Hochschule Ingolstadt, AI Motion Bavaria, Ingolstadt, Germany\medskip\\
	\affTwo Technical University Munich, Chair of Traffic Engineering and Control, Munich, Germany\medskip\\
	*Correspondence: Jeremias Gerner, jeremias.gerner@thi.de
}

\TIBtitle{Realistic and Scalable Floating Car Observer Detection in SUMO Derived from Co-Simulation}
\TIBsubtitle{}
% \TIBbundlename[ConfAbbrev]{Full conference name}
% \TIBconferencesession{Session Name}%ignored for journal articles
% \TIBdoi{to be filled in}
% \TIBpublisheddate{to be filled in}
\TIBabstract{%
Floating Car Observers (FCOs) are perception-equipped, connected vehicles that share object-level detections of surrounding traffic participants via V2X communication \cite{zhang2023novel,ma2021high,florin2022real}. By extending classical Floating Car Data (FCD) with onboard sensor outputs, FCOs enable mobile, infrastructure-independent traffic state estimation with broad spatial coverage. FCO-derived observations are therefore highly relevant for ITS applications where a complete and accurate picture of the current traffic state is essential. A key challenge in FCO research is the simulation-based development and evaluation of algorithms that rely on realistic detection inputs. Microscopic traffic simulators like SUMO operate at a high level of abstraction and do not natively model sensor physics or the stochastic behavior of modern learning-based perception systems. Existing approaches approximate FCO detectability via simple distance thresholds \cite{zhang2023novel} or 2D ray tracing \cite{Ilic2024raytracing,DuensingHoyer2025FCO}, producing simplified detections not grounded in real-world sensor capabilities. Furthermore, they assign binary detected/undetected labels and fail to capture the uncertainty, position error, and sensor-dependent characteristics of real 3D object detectors.

To address this gap, we present a scalable neural network-based observation model \cite{gerner2023sumodetector,IV2024} that emulates the outputs of state-of-the-art 3D object detectors directly within SUMO. Training data is generated through a synchronized SUMO--CARLA \cite{CARLA2017} co-simulation, in which observer vehicles are equipped with virtual camera and LiDAR sensors following the NuScenes sensor setup \cite{nuscenes}. Detections are produced by fine-tuned 3D detection models covering LiDAR-only, camera-only, and multimodal configurations, and capture realistic imperfections including localization error and missed detections. A lightweight neural network, trained on abstract scene representations natively available in SUMO---capturing surrounding traffic participants and infrastructure geometry---then reproduces these detection outcomes at runtime without requiring high-fidelity rendering or costly detector inference. This enables large-scale SUMO simulations with realistic FCO perception: detection evaluation runs in under 10\,ms per FCO, on par with 2D ray tracing, while achieving over 93\,\% detection accuracy compared to CARLA ground truth.

At the conference, we present via a poster and live demonstration how the detection module is trained, illustrating the full process from co-simulation to a scalable detection module. The detection module and trained neural network weights are shared via an open-source repository (\url{https://github.com/urbanAIthi/fco}, shows a first version) to enable other researchers to use the detection module for Floating Car Observer and general detection-based use cases in SUMO.%
}
\TIBkeywords{Floating Car Observers, Traffic State Estimation, SUMO, Co-Simulation, Neural Network, Extended Floating Car Data}
\addbibresource{local.bib}

% ----------------
\usepackage{xfrac}
% \usepackage{put your packages here}
% ----------------

% Suppress header with conference/author info
\chead{}

% Override maketitle to remove conference/DOI/license/date header
\makeatletter
\renewcommand{\maketitle}{%
    \begin{center}
        \LARGE\textbf{\contribtitle}\\\smallskip
        \Large{\contribsubtitle}
    \end{center}
    \begin{center}
    \normalsize{\@author}
    \end{center}
    \begin{center}\small
        \affiliations
    \end{center}
    \textbf{\large Abstract.} {\large\thisabstract}\bigskip\\
    \noindent\textbf{\large Keywords:} {\large\keywords}
    \thispagestyle{plain}
    \fontsize{12pt}{14pt}\selectfont
}
\makeatother

\begin{document}
\maketitle

\printbibliography[heading=references]

\end{document}
